---
title: "Gun-Murders"
author: "Vishal Jotwani"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This is a report on 2010 gun murder rates obtained from FBI reports. The original data was obtained from [this Wikipedia page](https://en.wikipedia.org/wiki/Murder_in_the_United_States_by_state).

We are going to use the following libraries:

```{r loading-libs, message=FALSE}
library(tidyverse)
library(ggrepel)
library(ggthemes)
library(ggrepel)
```
  
and load the data we already wrangled:

```{r}
load("rdas/murders.rda")
```

## Computing average murder rate in USA

```{r avg}
r <- murders %>% summarize(rate = sum(total) / sum(population) * 10 ^ 6) %>% pull(rate)
```

## Murder rate by Population in millions for every State 

We note the large state to state variability by generating scatterplot showing the murder rate by population in millions.

```{r murder-rate-by-population, echo=FALSE}
p <- murders %>% ggplot() 

p + geom_point(aes(x = population / 10 ^ 6, y = total, col = region), size = 3) + 
  geom_text_repel(aes(x = population / 10 ^ 6, y = total, label = abb), nudge_x = 0.01) +
  geom_abline(intercept = log10(r),lty = 2, color = "darkgrey") +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10") +
  xlab("Populations in millions (log scale)") +
  ylab("Total number of murders (log scale)") +
  ggtitle("US Gun Murders in 2010") +
  scale_color_discrete(name = "Region") +
  theme_few()
```

We note that for south region most of the states are above country's average murder rate.